Supplemental Materials for ``Universals of word order reflect optimization of grammars for efficient communication''

### 2019-10-27 writeup/pnas-word-order-si.tex
code

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, ordered by corpus size, measured by the number of sentences in the training partition, from largest (Czech) to smallest (Irish). Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars. All data are $z$-scored.}\label{fig:pareto-per-lang}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-raw.pdf}
	\caption[Predictability and Parseability]{Raw numerical values estimated for Predictability (negative surprisal), and negative syntactic ambiguity $-H[T|U]$, before $z$-scoring. For more meaningful comparison, both quantities are normalized by the number of words in the corpus, i.e., we plot per-word negative surprisal and per-word difficulty in recovering syntactic structures.  Note that the negative syntactic ambiguity $-H[T|U]$ equals parseability $I[T,U] = H[T] - H[T|U]$ up to a per-language constant $H[T]$, which we do not attempt to estimate. Further note that we use different scales in the different panels. }\label{fig:pareto-per-lang-raw}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/pareto-plane-perLanguage-arrows-smoothed-halfspace-untransformed.pdf}
	\caption[]{Per-language results as in Figure~\ref{fig:pareto-per-lang-raw}, representing the distribution of baseline grammars with a Kernel density estimate. We add lines connecting those points that have the same efficiency as the real grammar at $\lambda = 0.0$ (dotted) and $\lambda = 0.9$ (dashed). Points to the bottom/left of these line have lower efficiency than the real grammar, at the given value of $\lambda$.}\label{fig:lambda-halfplane-09}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-WithN.pdf}
	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, ordered by corpus size, measured by the number of sentences in the training partition, from largest (Czech) to smallest (Irish). Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars. All data are $z$-scored.}\label{fig:pareto-per-lang}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-raw-WithN.pdf}
	\caption[Predictability and Parseability]{Raw numerical values estimated for Predictability (negative surprisal), and negative syntactic ambiguity $-H[T|U]$, before $z$-scoring. For more meaningful comparison, both quantities are normalized by the number of words in the corpus, i.e., we plot per-word negative surprisal and per-word difficulty in recovering syntactic structures.  Note that the negative syntactic ambiguity $-H[T|U]$ equals parseability $I[T,U] = H[T] - H[T|U]$ up to a per-language constant $H[T]$, which we do not attempt to estimate. Further note that we use different scales in the different panels. }\label{fig:pareto-per-lang-raw}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/pareto-plane-perLanguage-arrows-smoothed-halfspace-untransformed.pdf}
	\caption[]{Per-language results as in Figure~\ref{fig:pareto-per-lang-raw}, representing the distribution of baseline grammars with a kernel density estimate. We add lines connecting those points that have the same efficiency as the real grammar at $\lambda = 0.0$ (dotted) and $\lambda = 0.9$ (dashed). Points to the bottom/left of these line have lower efficiency than the real grammar, at the given value of $\lambda$.}\label{fig:lambda-halfplane-09}
\end{figure}
